Detecting changes in dispersion in COVID-19 incidence time series using a negative binomial model

### 2023-05-20 noise_covid_v2.Rmd
debug all

plot(populations$population[which(sapply(ch_pts, length) != 0)], log(unlist(mean_pw)), xlab="Log Population Size", ylab=" Log Mean PW Statistic", col=2, xlim=c(0,1000000))
```
Methods designed specifically for count data show a similar dependence on population size. The cpt.meanvar function (changepoint package) is designed to detect changes in mean and variance in count data. The number of changepoints identified for a time series increases with its population size.

@@

state <- "FL"
source("code/wald_counties.R")
```
```{r, fig.cap = "P-values resulting from testing whether the dispersion parameter is smaller after Thanksgiving in each county in New York (LRT)"}
par(mfrow = c(1,2))
state <- "NY"
source("code/lrt_counties.R")
NY_sig <- (which(pvals < 0.05))

state <- "FL"
source("code/lrt_counties.R")
FL_sig <- (which(pvals < 0.05))
```

```{r}
signif_index <- c(NY_sig, FL_sig)

@@

source("code/vis_signific_state.R")
```
```{r}
# disp_stats <- c()
# for (i in 1:dim(new_cases_subset[state_index])[1]){
#   Y <- new_cases_subset[i,]
#   spline_mod.1.pois <- spl_fit_pois(Y = Y, population=ny_fl_df_t$populations[i], inds=1:(length(Y)/2))
#   disp_stats <- c(disp_stats, spline_mod.1.pois$deviance/spline_mod.1.pois$df.residual)
# }
```

```{r}
# library(tigris)
# p <- round(fl_df_t$p, digits=1)
# p[is.na(p)] <- "na"
# fl$p <- p 
# 
# ggplot(fl) +
#   geom_sf(aes(fill=p))
```

@@

>>>>

plot(populations$population[which(sapply(ch_pts, length) != 0)], log(unlist(mean_pw)), xlab="Log Population Size", ylab=" Log Mean PW Statistic", col=2, xlim=c(0, 1000000))
```
Methods designed specifically for count data show a similar dependence on population size. The cpt.meanvar function (changepoint package) is designed to detect changes in mean and variance in count data. The number of changepoints identified for a time series increases with its population size.

@@

state <- "FL"
source("code/wald_counties.R")
```
```{r}
library(tigris)
fl <- counties("Florida")
load("data/W_pvalsFL.Rdata")
p <- round(pvals, digits=1)
p[is.na(p)] <- "na"
fl$p <- p

ggplot(fl) +
  geom_sf(aes(fill = p)) 
```

```{r, fig.cap = "P-values resulting from testing whether the dispersion parameter is smaller after Thanksgiving in each county in New York (LRT)"}
par(mfrow = c(1,2))
state <- "NY"
source("code/lrt_counties.R")
NY_sig <- (which(pvals < 0.05))

state <- "FL"
source("code/lrt_counties.R")
FL_sig <- (which(pvals < 0.05))
```
Note that in these visualizations, an NB2 fit would be:
A Poisson fit would be:
```{r}
signif_index <- c(NY_sig, FL_sig)

@@

source("code/vis_signific_state.R")
```

@@

```{r}
load("data/pvals_sim_Wald.Rdata")
mean(pvals[which(curve_parms$theta1==curve_parms$theta2)], col=1)
mean(pvals[which(curve_parms$theta1!=curve_parms$theta2)], col=2)

load("data/pvals_sim_LRT.Rdata")
mean(pvals[which(curve_parms$theta1==curve_parms$theta2)], col=1)
mean(pvals[which(curve_parms$theta1!=curve_parms$theta2)], col=2)

```
